---
title: "mp4 title"
author: "Mathena Abramson"
date: "December 1, 2017"
output:
  html_document:
    code_folding: hide
---
```{r, message = FALSE, warning = FALSE}
library(mdsr)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
```

Are sequels today more frequent than in movie’s past or are we just simply more aware of it all? We set out to investigate how the use of sequels within the film industry has changed. We hypothesized that the frequency of produced sequels would increase overtime. In order to examine this trend, we utilized information from the imdb data set and joined the datatables titles and `movie_link`. Using this data, we selected sequels by filtering films that had been coded by imdb as full-length movies following original films.To select the movies that were sequels we filtered for when `link_type_id` = 1. This allowed us to see movies that follow another. Additionally, we selected films produced between 1900 and 2020. As a result, we created a line graph showing the number of sequels produced over time.

```{r, message = FALSE, warning = FALSE}
sqls_data <- db %>%
  dbGetQuery("SELECT COUNT(DISTINCT t.id) AS count, title, kind_id, link_type_id, production_year
FROM title t
JOIN movie_link ON t.id = linked_movie_id
WHERE kind_id = 1 AND link_type_id = 1
GROUP BY production_year;")
```

```{r, message = FALSE, warning = FALSE}
sqls_plot <- sqls_data %>%
  ggplot(aes(x = production_year, y = count)) +
	geom_line(size = 1) +
  geom_area(fill = "#d3a3fa", alpha = .5)
```

```{r, message = FALSE, warning = FALSE}
sqls_plot +
  scale_x_continuous(expand = c(0,0),
                   breaks = c(1900, 1910, 1920, 1930, 1940, 1950,
                              1960, 1970, 1980, 1990, 2000, 2010,
                              2020),
                   labels = c("1900", "1910", "1920", "1930",
                              "1940", "1950", "1960", "1970",
                              "1980", "1990", "2000", "2010",
                              "2020"),
                   limits = c(1900, 2020)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        panel.border=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_rect(fill = "#f5f5f5"),
        panel.background=element_blank()) +
  labs(title = "Trend of Movie Sequels Over Time",
       x = "Production Year",
       y = "Number of Sequels",
       caption = "(Data from IMDb)")
```

Looking at this graphic, sequels spiked in popularity between 2005-2017. The resulting trend supports our initial hypothesis that the frequency of sequels increased overtime.This could be explained by the release of several popular sequels in movie franchises such as: Harry Potter, Shrek, Toy Story, Saw, Star Wars, Pirates of the Caribbean, and The Ocean's Trilogy.  However, there are also spikes in the number of sequels between 1910-1920 and 1922-1935. These spikes suggest that while sequels have become more popular in recent years, they have a history of being popular within the film industry. During the early 20th-century, this trend might be explained by the rise of popular movie stars and their repeated appearances in movie sequels. For example, Rudolph Valentino’s performance in The Sheik (1921) was a tremendous hit and subsequently prompted the sequel The Son of the Sheik (1926). Alternatively, the spike in sequels from the 1910-1930s could be explained by IMDb’s operationalization of sequels and follow-up  films. For example, the link id we selected (follows)  defined sequels in a multitude of ways which differ from everyday consumers’ conceptions of movie sequels. For instance, IMDb defines film adaptations of books sequels as reboots or follow-up films. As a result, additional research needs to be done on IMDb’s operationalization of variables in order to determine whether the trend depicted in the graphic accurately reflects the film industry.

#Git Repository
https://github.com/m-abramson/mp4
